### 2018-02-01 paper/conclusion.tex
First draft

\section{Conclusion}

In this paper, a batched Bayesian optimization algorithm is proposed for the
automation of analog circuit design. the parallelization is achieved via
multi-objective ensemble of acquisition function. In each iteration, the
candidate points are sampled from the Pareto front of multiple acquisition
functions. We compared the proposed MACE algorithm using analytical benchmark
functions and real-world circuits, it is shown that the MACE algorithm is
competitive compared with state-of-the-art method listed in the paper.