### 2018-02-03 paper/abstract.tex
Paper update

\begin{abstract} 
    The Bayesian optimization algorithm has recently shown promising
    performance for the automated design analog integrated circuits. However,
    the sequential property of Bayesian optimization limits its further
    application. In this paper, a batched bayesian optimization algorithm is
    proposed, the parallelization is realized via a multi-objective ensemble of
    acquisition functions. In each iteration, multiple acquisition functions
    are selected, the multi-objective optimization is performed to search for
    the Pareto front of the acquisition functions, candidate points are then
    sampled from the Pareto front. The proposed method is compared with several
    state-of-the-art batched Bayesian optimization algorithms using analytical
    benchmark functions and real-world analog circuits; the experimental
    results show that the proposed method is competitive compared to the
    state-of-the-art algorithms.
\end{abstract}