### 2018-02-07 paper/abstract.tex
revise abstract

\begin{abstract}
    Bayesian optimization methods are promising for the optimizations of black-box functions that are expensive to evaluate. In this paper, a novel batch Bayesian optimization approach is proposed. The parallelization is realized via a multi-objective ensemble of multipe
    acquisition functions. In each iteration, the multi-objective optimization of the multiple acquisition functions is performed to search for
    the Pareto front of the acquisition functions. The batches of inputs are then selected from the Pareto front. The Pareto front represents the best trade-off between the multiple acquisition functions. Such a policy for batched Bayesian optimization can significantly improve the efficiency of optimization. The proposed method is compared with several
    state-of-the-art batch Bayesian optimization algorithms using analytical
    benchmark functions and real-world analog circuits. The experimental
    results show that the proposed method is competitive compared with the
    state-of-the-art algorithms.
\end{abstract}