### 2018-02-08 paper/conclusion.tex
minor revision of experimental results and conclusion

\section{Conclusion}

In this paper, a batch Bayesian optimization algorithm is proposed for the
automation of analog circuit design. The parallelization is achieved via
the multi-objective ensemble of acquisition functions. In each iteration, the
candidate points are sampled from the Pareto front of multiple acquisition
functions. We compared the proposed MACE algorithm using analytical benchmark
functions and real-world circuits, it is shown that the MACE algorithm is
competitive compared with the state-of-the-art methods listed in the paper.