### 2018-06-06 paper/mo_description.tex
Supplementary

\section{Description and Demonstration of the DEMO Multi-objective Optimization Algorithm}