### 2018-06-06 paper/mo_description.tex
Revise the supplementary materials

\section{Description and Demonstration of the DEMO Multi-objective Optimization Algorithm}
% TODO: DE operations
\paragraph{Evolutionary optimization}
\paragraph{DE operations}
\paragraph{Non dominated sorting}
\paragraph{DEMO}
\paragraph{parameters of DEMO}